 \section{The electronic wave function of the hydrogen molecular ion\\ \small{(%Pathway: AMA; 
 Pre-requisite: Quantum theory)}}
%\item
%\underline{\bf The electronic wave function of the hydrogen molecular ion} 
\setcounter{equation}{0}
\begin{enumerate}\setlength{\itemsep}{0pt}
\item[(a)] The stationary states of the electron in the hydrogen atom are
described by the wave function $\psi ({\bf r})$ and energy $E$, which are
found from the Schr\"odinger equation
\begin{equation}\label{eq:Sch}
\left [ - \frac{\hbar^2}{2m} \nabla^2 ~ - ~ \frac{e^2}{r} \right ]
\psi(\mbox{\bf r}) ~ = ~ E \psi(\mbox{\bf r})
\end{equation}
where $m$ is the electron mass, $-e$ is its charge, $\hbar $ is the reduced
Planck's constant and the proton is placed at the origin $r=0$.
Equation (\ref{eq:Sch}) separates when spherical polar coordinates
($r$, $\theta$, $\phi$) are used. Thus if we write $\psi ({\bf r})=
R(r)\Theta(\theta)\Phi(\phi)$, we find separate differential equations for
each of the functions $R(r)$, $\Theta(\theta)$ and $\Phi(\phi)$.\\[6pt]
Derive these equations for $R(r)$, $\Theta(\theta)$ and $\Phi(\phi)$,
and explain how they are solved.
\item[(b)] Now consider an electron moving in the field of {\em two} nuclei,
each with charge $e$, at fixed points $A$ and $B$. Let $r_A$ and $r_B$ be the
distance of the electron from $A$ and $B$ respectively, and let $R \equiv |AB|$.
\begin{itemize}
\item What is the equation, analogous to equation (\ref{eq:Sch}), that is
satisfied by the electronic wave function $\psi$({\bf r})?
\item Show that this differential equation separates when {\em prolate
spheroidal coordinates} ($\lambda, \mu, \phi$) are used, where
$\lambda \equiv (r_A+r_B)/R$, $\mu \equiv (r_A-r_B)/R$ and $\phi$ is now the
angle of rotation about $AB$.
\item If $\psi$({\bf r}) = $\Lambda(\lambda)M(\mu)\Phi(\phi)$, find the
equations satisfied by $\Lambda(\lambda)$, $M(\mu)$ and $\Phi(\phi)$.
\item Discuss a strategy for solving these equations.
\end{itemize}
\end{enumerate}
{\bf References}\\[3pt]
L.~I.~Schiff,  {\em Quantum Mechanics}, \S14 and \S16. \\[3pt]
M.~Abramowitz and I.~A.~Stegun,  {\em Handbook of Mathematical Functions},
Ch.~21. \\[3pt]
D.~R.~Bates and R.~H.~G.~Reid, {\em Adv. At. Mol. Phys.} {\bf 4}, p.~13.

